Numerical experiment investigating the influence of drag on trajectory patterns of floating macroalgae

### 2022-08-31 thesis_chapter_3.Rmd
uploaded updated analysis files

---
title: 'Numerical experiment investigating the influence of drag on trajectory patterns of floating macroalgae'
author: 
- R Coppin^[Department of Biodiversity and Conservation Biology, University of the Western Cape, Private Bag X17, Bellville 7535, South Africa], C Rautenbach^[National Institute of Water and Atmospheric Research, Hamilton: Gate 10 Silverdale Road Hillcrest, 3216], AJ Smit^[South African Environmental Observation Network, Elwandle Coastal Node, Port Elizabeth, South Africa]
output:
  html_document:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
  word_document:
    reference_docx: markdown/wiley_template.docx
bibliography: references.bib
csl: markdown/limnologyandoceanography.csl
abstract: "**Abstract**: Ocean currents act as an important dispersal mechanism of material on the ocean surface, both natural and anthropogenic. Macroalgae are one of the most important natural dispersal vectors  of marine organisms and are regarded as the 'tumble-weed of the ocean'. Despite many studies that exist on the topic, the relative role of wind and surface currents in influencing the trajectory of macroalgal dispersal is still unknown. Past studies focus on kelp-rafts, which can vary greatly in size, making it difficult to form a consenses as to the relative role of wind versus surface currents. In addition, these studies have not considered surface-area in relation to drag characteristics of the macroalgae, both of which have been shown to play a role in the trajectory and accumulation of other flotsam. Advances in Lagrangian trajectory modeling have been rapid in recent years and allow the use of numerical experiments to investigate trajectories of flotsam in the ocean. This study aimed to shed light on the relative role of wind versus surface currents, and the role of drag in macrolagal trajectory and accumulation. We focused on solitary kelp individuals. This was achieved by comparing simulations of virtual kelp 'particles' which incorporate drag with that of purely Lagrangian particles. The results showed that the virtual kelp particles do follow the ocean currents in general, but that the inclusion of drag into the scenarios causes accumulation in eddys within the study domain. The results also point to the importance of sub-mesoscale features, such as eddys, as important surface accumulation zones for solitary floating macroalage."
---

@@

Passive dispersal mechanisms of marine taxa include a range of natural and anthropogenic objects that float in the ocean, and macroalgae are regarded as one of the most important of these. Floating kelps, sometimes referred to as the ‘tumble-weed of the ocean,’ are very important in this regard [@edgar1987; @bushing1994; @norton1992; @helmuth1994; @holmquist1994; @smith2002; @McCormick2008]. Kelps also function as allochthonous materials to near and distant ecosystems and in this manner support the ecological function in a range of adjacent habitats[@cartraud2021; @colombini2003; @Filbee-Dexter2018-wp; @zuercher2019; @Krumhansl2012; @vanderklift2008; @bustamante1995].

Some macroalgal species are negatively buoyant and sink to the seafloor when detached from the substrate, while other species have air-filled pneumatocysts or stipes which allow the plants to reach the surface where light is more abundant. The positively buoyant pneumatocysts cause plants to float to the surface when dislodged from the substratum. The structure and number of pneumatocysts vary among species. For example, species belonging to the genera *Macrocystis*, *Sargassum*, *Ascophyllum*, and *Fucus* have thalli with many small pneumatocysts, while other species such as *Nereocystis luetkeana*, *Pelagophycus porra*, *Ecklonia radiata*, and *Ecklonia maxima* have single, large pneumatocysts[@dayton1985; @smith2002; @thiel2005; @graiff2016; @batista2018]. In some cases, the stipe itself may be air-filled, such as with *E. maxima* and *E. radiata*. Although there have been reports of floating species of Chlorophyta and Rhodophyta, Phaeophyceae species are the most reported forms of floating algae. This is most likely because the green and red species reported to float are not actually positively buoyant, but instead are kept at the surface by gas trapped in-between or in their thalli [@dromgoole1982; @back2000]. The giant kelp *Macrocystis pyrifera* [@hobday2000; @macaya2005; @graiff2016; @batista2018], the bull kelp *Durvillaea antarctica* [@tala2017; @wichmann2012; @batista2018], and pelagic *Sargassum* species [@putman2020; @brooks2019; @brooks2018; @wang2019] have been the focus of much of the research regarding spatial and temporal dispersal patterns, the ecological role of rafting, marine connectivity and raft-time.

Past research points to macroalgal trajectories being largely determined by prevailing wind conditions and surface currents [@thiel2005; @hobday2000; @fraser2011; @rothausler2011; @rothausler2011]. Although ocean currents are regarded as the primary influence, the relative importance of wind versus surface current is still unknown, although the role of wind has been recognized as important in several studies. For example, a study by @harrold1989 investigated the seasonal trajectories of radio-transmitter tagged *M. pyrifera* in nearshore Monterey Bay. The results showed that kelp rafts, with little surface area exposed to the wind, were largely driven by a combination of wind and surface currents; however, the relative importance of wind and was not clear. In addition, the trajectories of tagged kelp were more consistent with the formation of eddies during winter. Studies have also identified wind as an important mechanism of dispersal in wind dominated ocean systems. For example, in the sub-Antarctic latitudes the West Wind Drift causes continuous unidirectional surface flow and is regarded as an important mechanism for dispersal of floating bull kelp *Durvillaea antarctica* [@waters2008; @smith2002; @nikula2010]. Other studies have used genetic approaches to determine macroalgae raft trajectory characteristics by inferring source location from genetically similar populations [@nikula2013]. For example, a study by @fraser2011 on the rafting capabilities of *D. antarctica* used a combination of population genetics and relative age estimate of goose barnacles attached to the raft. The presence of goose barnacles suggests a long rafting time as these species have a slow growth rate. Genetic analyses showed these species can raft up to $\sim$ 390 km from their local origin. The authors suggested that wind and water-movement were the primary influences of trajectory; however, this was only inferred from the genetic results and local climatology data [@fraser2011]. Research has shown that buoyancy and drag also play a role in determining the trajectory and rate of transport for surface floating material [@onink2019; @shen2019; @dasaro2018; @gates2018; @beron2016; @trinanes2016; @furnans2008]. The 'sail' area of an object floating at sea is the surface area exposed to the wind which results in air form drag, while the area of the object below the surface of the water is exposed to surface currents which result in hydrodynamic-form drag [@breivik2011; @griffin2017]. Work by the maritime industry has shown that the size and shape of a vessel determine the relative importance of waves or wind as drivers of trajectory, as well as orientation of the object. If the length of the object is longer than the significant wave height then waves will be the primary driver of trajectory while the opposite is true for smaller objects where the effect of waves is regarded as negligible [@breivik2011; @griffin2017]. Drag is ultimately determined by the size and shape of the object which are properties that vary considerably with macroalgae species.

To accurately determine the trajectory of marine macroalgae, both water and wind drag need to be considered. @allen1999 and @breivik2011 have provided estimates of drag for various objects based on experimental work, which consisted of both direct and indirect methods of the Leeway model. However, this work did not consider biological material such as macroalgae. Although drag estimates do exist, these have been applied for macroalgae not detached from the substratum and are regarded as fixed-point estimates [@Denny1997; @Denny2002a; @Denny2002b; @Gaylord1997-ow; @Gaylord2008-qv]. Although various approaches have been used in the past to investigate floating macroalgal trajectories, very few studies have employed the use of Langrangian trajectory modeling. Furthermore, none of these studies using this modeling approach have considered macroalgal morphology cross-sectional area (*i.e.* shape) as an aspect of drag and ultimately trajectory. @brooks2019, @putman2018, and @putman2020 used Lagrangian approaches to investigate the effects of inertia, raft-size, and windage on macroalgae trajectory. For example, @brooks2019 used a custom growth model to estimate changes in biomass and ultimately radial size, while a customized Hybrid Coordinate Ocean Model (HYCOM) for the trajectory simulations was used. Their results showed that the trajectory of pelagic *Sargussum* was significantly influenced by inertia and the radial size of the rafts. @putman2020 investigated the effect of including windage in macroalgae trajectory simulations and showed that including *ad hoc* windage factors greatly improved accuracy regarding entrainment or expulsion from eddys when compared to tracked *Sargassum* mats in the ocean. Langrangian trajectory modeling is clearly a useful tool; however, the various models available often assume the object to be a spherical particle, which is obviously not the case for both biological and anthropogenic forms of marine debris in nature. The radial size included by @brooks2019 is an improvement but this approach does not consider the complex morphology on an individual level into account.

Studies have focused on kelp-rafts, not solitary floating kelp, and therefore are not able to investigate the role of eddys and vortices in the accumulation of individual kelp plants at the surface. Gyres, eddys, and converging surface currents have been identified as important in the accumulation and dispersal of plastics and other marine debris [@onink2019; @dasaro2018; @brach2018]. Investigations into  the dispersal of marine debris have also included the possible effects of inertia in accumulation of ocean drifter instruments in mesoscale features. For example, @beron2016 used a modeling approach to investigate the role of Eckman transport and inertia to explain the observed behavior of undrogued drifters in the ocean gyres. Undrogued drifter trajectory is affected by wind, unlike the drogue counter parts which makes the effects of wind negligible. The results showed that the combined effect inertia, not Eckman transport, played a significant role in the observed drifter behavior [@beron2016]. These authors also noted that the presence of other flotsam, along with the undrogued drifters in ocean gyres, suggests that aspects of inertia also play a role in accumulation of other marine debris [@beron2016]. The approach by @beron2016 has also been improved upon by @beron2019 but they also highlight the significant role of inertia in flotsam trajectory and accumulation.

In this study, a numerical approach is used to shed light on the role of hydrodynamic and direct wind drag on macroalgal drift trajectory using Lagrangian based trajectory simulations. We aim to determine the 'behavior' of surface drifting macroalgae from a coastal environment to the open ocean, which will inform future characterization studies. We will achieve this by comparing the trajectory outputs of virtual passive particles (no drag characteristics added) with kelp particles (drag characteristics added) of varying degrees of hydrodynamic and wind drag exposure. The different degrees of wave and wind exposure are meant to reflect the different buoyancy characteristics, as well as identify the role of magnitude of the drag components on overall trajectory. The modeling of single kelp offers the opportunity to study the role of eddys and vortices in accumulation points on the surface.

# Methods
## Study site

The region of interest for this study was the Cape Peninsula in the Western Cape (Figure 1A). Kommetjie, located on the Atlantic side of the Cape Peninsula (Figure 2B), was chosen as a release site. The hydrodynamics in the study region are characterized by a complex interaction of wind, sea waves and ocean swell, and are also influenced by larger-scale regional oceanic processes. The larger processes in the region can be attributed to the Agulhas Current (AC) and the Benguela Current (BC) [@lutjeharms2000;@lutjeharms2007;@garzoli1996;@hutchings2009;@rubio2009]. The AC is a component of the South-west Indian Ocean sub-gyre, flowing predominantly south-westward following the continental shelf edge and eventually retroflecting eastward back into the South Indian Ocean [@lutjeharms2000;@lutjeharms2007].

The region where the retroflection occurs is known as the Agulhas Retroflection and consists of high levels of mesoscale variability [@garzoli1996]. The variability results in shedding of rings and filaments known as the Agulhas Rings and have an average diameter of ~324 km [@lutjeharms1988;@lutjeharms2006; @lutjeharms2007]. Agulhas Rings are the main contributors to the Agulhas leakage which is the transport of warm water from the Indian Ocean to the South Atlantic Ocean [@garzoli1996; @lutjeharms2006; @beal2015]. The Agulhas leakage is also a source of meso-scale variability within the Benguela region as anticyclonic rings are shed and drift northwestward along the southern Benguela slope region [@rubio2009]. The Benguela region ---known as the Benguela Upwelling System (BUS)--- is characterized by strong coastal upwelling which is primarily driven by equatorward winds [@shannon1996;@hardman2003;@veitch2018; @veitch2017]. The BUS starts at 27$^\circ$S and extends southward to 35$^\circ$S [@shannon1996]. The region is also characterized by large amounts of mesoscale variability in the form of eddys and filaments [@blanke2002; @blanke2005; @rubio2009]. Inshore, the BUS is dominated by a northwestward flow, known as the Benguela Jet, because of topographical steering, wind stress and interactions with passing Agulhas Rings and eddys [@stramma1989; @veitch2010; @veitch2017; @veitch2018; @hutchings2009]. The stability of the jet varies seasonally with more stable flows in winter; in January, the jet begins to split between Cape Point and Cape Columbine [@veitch2018; @blanke2005; @blanke2002] which results in a weak narrow inshore branch that flows along the shelf edge and a strong flowing north-westwards offshore branch. The area is also characterized by high upwelling activity driven by strong westerly winds [@hutchings2009]. These winds are most intense during Austral summer (December-March) which favors upwelling conditions in the region, and also act to reinforce the flow of the Benguela Jet which occurs due to cold upwelling waters contributing to the across-shore density gradient [@ragoasha2019]. In Austral winter the wind direction is reversed which decreases upwelling favorable conditions, but the core of the Benguela Jet is maintained by the intrusion of warm Agulhas waters into the Benguela region over the shelf edge [@veitch2018; @hutchings2009; @ragoasha2019].

```{r map, echo=FALSE, message=FALSE, warning=FALSE, strip.white=TRUE, fig.cap="Figure 1: Map of the study domain and the main features pertaining to transport in the region. In panel A, the map of the western portion of South African coastline and the main currents is presented. The resulting eddys and filaments froming from the Agulhas retroflection are indicated, and the position of the study area is drawn in as a black rectangle. Detail of the study area is in panel B. Here the splitting of the Benguela Jet, eddy features, and the general flow within the study domain are presented. The release site of simulated kelp ‘particles’ is represented by a blue star."}

@@

@@

The model is based on parameterizations that have been used for previous studies investigating iceberg, capsized marine vessels and microplastics. It should be noted that most of the previous studies regarding macroalgal trajectories have investigated rafts and not solitary floating individuals. A kelp raft will vary in size and can range from a few meters across to large dense mats the size of a sports field. Some authors suggest that most macroalgae become entangled through ocean currents and not through the hydrodynamic forces that dislodge them. Large kelp rafts are not a feature around the coast of the Cape Peninsula; rather, solitary kelp is (pers. obs). Therefore, the trajectory of only solitary *E. maxima* individuals will be considered. We performed numerical calculations of particle trajectory with a Lagrangian method [@delandmeter2019], which can be described by the equation:

$$X(t+\Delta t) = X(t) + \int_{t}^{t+\Delta t} v(x, \tau)d \tau + \Delta X_b(t)$$
Here, $X$ is the three-dimensional position of a particle, $v(x,t)$ is the three-dimensional velocity field at the location in the ocean model at particle point in time, $t$ is time $X_b(t)$ is a change in position due to 'behavior' of the particle at a particular point in time. To investigate the effects of drag components of trajectory of floating macroalgae, this study incorporated drag as the custom behavior.

>>>>

---
title: 'Numerical experiment investigating the influence of drag on trajectory patterns of floating macroalgae'
author: "R Coppin^[Department of Biodiversity and Conservation Biology, University of the Western Cape, Private Bag X17, Bellville 7535, South Africa], C Rautenbach^[National Institute of Water and Atmospheric Research, Hamilton: Gate 10 Silverdale Road Hillcrest, 3216], AJ Smit^[Department of Biodiversity and Conservation Biology, University of the Western Cape, Private Bag X17, Bellville 7535, South Africa, South African Environmental Observation Network, Elwandle Coastal Node, Port Elizabeth, South Africa]"
output:
  pdf_document:
    latex_engine: xelatex 
    toc: no
  html_document:
  word_document:
    reference_docx: markdown/frontiers_template.docx
bibliography: references.bib
csl: markdown/elsevier-harvard2.csl
abstract: "Ocean currents act as an important dispersal mechanism of natural and anthropogenic material on the ocean surface. Macroalgae are one of the most important natural dispersal vectors of marine organisms and are regarded as the ‘tumble-weed of the ocean.’ Despite many studies on the topic, the relative role of wind and surface currents in influencing the trajectory of macroalgal dispersal is still uncertain. Past studies focused on kelp-rafts, which can vary greatly in size, making it difficult to form a consensus as to the relative role of wind versus surface currents. In addition, these studies have not considered surface-area in relation to drag characteristics of the macroalgae, both of which have been shown to play a role in the trajectory and accumulation of flotsam. Advances in Lagrangian trajectory modeling have been rapid in recent years and allow the use of numerical experiments to investigate trajectories of flotsam in the ocean. This study aimed to shed light on the relative role of wind versus surface currents, and the role of drag in macroalgal trajectory and accumulation. We focused on solitary kelp plants. This was achieved by comparing simulations of virtual kelp ‘particles’ which incorporate drag with that of purely Lagrangian particles (no drag). Geometric representations for each morphological characteristic was used to calculate the overall surface area used in the calculation of the relevant drag components (hydrodynamic and wind) for each scenario (submerged and partially submerged) and flow-field combinations (currents only, currents with wind, currents with Stokes drift, currents with wind and Stokes drift). The results showed that the virtual kelp particles do follow the ocean currents in general, but that the inclusion of drag into the scenarios causes accumulation in eddies within the study domain."
nocite:  |
  @Algaebase
  @Worms
header-includes:
  - \usepackage[left]{lineno}
  - \linenumbers
---

@@

Passive dispersal vectors of marine taxa include a range of natural and anthropogenic objects that float in the ocean, and kelp are regarded as one of the most important of these [@edgar1987;@norton1992;@bushing1994;@holmquist1994; @helmuth1994;@smith2002; @McCormick2008]. Floating macroalgae are regarded as important transport mechanisms for marine invertebrates which remain attached to the plant when dislodged from the substratum [@edgar1987;@norton1992;@bushing1994;@holmquist1994; @helmuth1994;@smith2002; @McCormick2008]. For example a study by @holmquist1994 investigated the long-distance dispersal of the bivalve *Gaimardia trapesina* Lamarck via drifting kelp. The authors used observational techniques to quantify kelp-rafts along with haphazard sampling of the kelp-rafts observed. Genetic samples were taken from the target species for each of the haphazardly sampled kelp-rafts for comparison amongst populations. The results suggest that kelp-rafting, along with the local wind patterns, most likely played a key role in the dispersal of  *G. trapesina* between genetically similar but spatially disconnected populations. Kelps ecological role in terms of organic subsidy is not limited to the local environment, and instead plays an important role in organic input into deep, offshore ecosystems [@bustamante1995; @colombini2003;@vanderklift2008;@Krumhansl2012; @filbee2018;@zuercher2019;@cartraud2021]. For example, a study by @filbee2018, investigated the allochthonous input of kelp detritus into deep fjords within northern Norway. The authors used a combination of video surveys, biomass estimates and tagging of detrital material to characterise kelp detritus within shallow, sub-tidal and deep fjord systems. The results showed that the organic subsidy into the different systems was linked to loss of biomass from kelp forests during times of high wave exposure. Furthermore, the results showed that kelp detritus is an important organic subsidy for distant ecosystems, suggesting resource coupling between nearshore kelp forests and offshore systems [@filbee2018]. The formation of kelp rafts also act as floating refuges for marine organisms as well as natural fish aggregation devices. Work by @witherington2012 showed the reliance of young sea turtles on drifting pelagic Sargassum rafts as a source of food and refuge from marine predators, and work by @kingsford1992 showed the reliance of some fish larvae (*Paralabrax clathratus* Girard, *Chromis punctipinnis* Cooper and Sebastes spp) on kelp-rafts for habitat complexity and transport to offshore areas.

Some macroalgal species are negatively buoyant and sink to the seafloor when detached from the substrate, while other species have air-filled pneumatocysts or stipes which allow the plants to remain at the surface where light is more abundant, and in turn cause plants to float to the surface when detached. The structure and number of pneumatocysts vary among species. For example, species belonging to the genera Macrocystis, Sargassum, Ascophyllum, and Fucus have thalli with many small pneumatocysts, while other species such as *Nereocystis luetkeana* (K.Mertens) Postels & Ruprecht, *Pelagophycus porra* (Léman) Setchell, *Ecklonia radiata* (C.Agardh) J.Agardh, and *Ecklonia maxima* (Osbeck) Papenfuss have single, large pneumatocysts [@dayton1985; @smith2002;@thiel2005; @graiff2016; @batista2018]. In some cases, the stipe itself may be air-filled, such as with *E. maxima* and *E. radiata*. Although there have been reports of floating species of Chlorophyta and Rhodophyta, Phaeophyceae species are the most reported forms of floating algae. This is most likely because the green and red species reported to float are not actually positively buoyant, but instead are kept at the surface by gas trapped in-between or in their thalli [@dromgoole1982; @back2000]. The giant kelp *Macrocystis pyrifera* (Linnaeus) C.Agardh [@helmuth1994; @kingsford1995;@hobday2000;@macaya2005; @graiff2016;@batista2018], the bull kelp *Durvillaea antarctica* (Chamisso) Hariot [@smith2002; @collins2010; @wichmann2012; @tala2013; @tala2017; @saunders2014; @batista2018], and pelagic Sargassum species [@brooks2018; @brooks2019; @putman2018; @putman2020; @wang2019] have been the focus of much of the research regarding spatial and temporal dispersal patterns, the ecological role of rafting, marine connectivity and raft-time.

Past research points to macroalgal trajectories being largely determined by  a combination of prevailing wind conditions and surface currents [@hobday2000; @thiel2005; @rothausler2011;@fraser2011]. Although ocean currents are regarded as the primary influence, the relative importance of wind versus surface current is still unknown, although the role of wind has been recognized as important in several studies. For example, a study by @harrold1989 investigated the seasonal trajectories of radio-transmitter tagged *M. pyrifera* in nearshore Monterey Bay. The results showed that kelp rafts, with little surface area exposed to the wind, were largely driven by a combination of wind and surface currents; however, the relative importance of wind was not clear. This work also showed that the trajectories of tagged kelp were congruent with the formation of eddies during winter. Other studies have used genetic approaches to determine macroalgal raft trajectory characteristics by inferring source location from genetically similar populations. For example a study by @nikula2013 used genetic markers (mtDNA and AFLP) of invertebrate species to show that in the sub-Antarctic latitudes the West Wind Drift causes continuous unidirectional surface flow and is regarded as an important mechanism for dispersal of floating bull kelp *D. antarctica*. Another study by @fraser2011 on the rafting capabilities of *D. antarctica* used a combination of population genetics and relative age estimate of goose barnacles attached to kelp-rafts. The presence of goose barnacles suggests a long rafting time as these species have a slow growth rate. Genetic analyses showed these species can raft up to 390 km from their local origin. The authors suggested that wind and surface currents were the primary influences of trajectory; however, this was inferred from the genetic analyses and local climatology data [@fraser2011].

Besides wind and surface currents, waves are also an important ocean feature and work by the maritime industry has shown that the size, shape, and orientation of the object of a vessel determine the relative importance of waves or wind as drivers of trajectory [@breivik2011; @griffin2016]. In general, if the length of the object is longer than the significant wave height then waves will be the primary driver of trajectory while the opposite is true for shorter objects where the effect of waves is regarded as negligible [@breivik2011; @griffin2016]. Stokes drift is a wave-driven process and previous work has identified this as an aspect that needs to be considered when investigating dispersal of macroalgal species. Stokes drift is the net drift velocity of a particle floating at the ocean surface in the direction of wave propagation due to the differences between the velocity of the fluid particle and flow-field [@bremer2018]. More generally, the particle floating at the surface moves slightly faster in the direction of the wave as a result of the oscillatory motion of fluid particles within a wave. Stokes drift can significantly influence the trajectory of a particle or object floating at the surface of the ocean. For example, work by @fraser2018 included Stokes drift into the advection scheme for the Lagrangian model used to investigate the dispersal potential of the southern bull kelp *D. antarctica* to reach Antarctica. The results showed that the inclusion of wave-driven Stokes drift and mesoscale eddies suggest frequent incursions into Antarctica, which the authors attributed to storm-forced surface waves, further highlighting the role of wave-driven processes in the open ocean.

Research has shown that buoyancy and drag also play a role in determining the trajectory and rate of transport for surface floating material [@furnans2008; @beron2016; @trinanes2016; @dasaro2018; @gates2018; @onink2019; @shen2019].  Therefore, to  accurately determine the trajectory of marine kelp, both water and wind drag need to be considered. Drag is the force an object experiences when moving through a fluid which is proportional to the velocity of the object [@Vogel2020]. The fluid viscosity and the friction from the surface of the object results in the overall drag component, as well as the surface area and shape of the object [@Vogel2020]. If the object is partially submerged, as a result of buoyancy forces, then drag will be a combination of drag from the air and fluid [@Vogel2020]. The ‘sail’ area of an object floating at sea is the surface area exposed to the wind (air) which results in air-form drag, while the area of the object below the surface of the water is exposed to surface currents which result in hydrodynamic-form drag [@breivik2011; @griffin2017]. @allen1999 and @breivik2011 have provided estimates of drag for various objects based on experimental work using both direct and indirect methods of the Leeway model, However, this work did not consider biological material such as macroalgae. Therefore, no drag estimates exist for macroalgae floating at the surface of the ocean and alternative approaches should be used to account for the drag components. Although drag estimates do exist, these have been applied for kelp not detached from the substratum and are regarded as fixed-point estimates [@Denny1997; @Gaylord1997-ow; @Denny2002a; @Denny2002b;@Gaylord1997-ow].

Although various approaches have been used in the past to investigate floating macroalgal trajectories, very few studies have employed the use of Langrangian trajectory modeling. Works by @brooks2019, @putman2018 and @putman2020, used Lagrangian approaches to investigate the effects of inertia, raft-size, and windage (drag from the wind) on kelp trajectory. For example, @brooks2019 used a custom growth model to estimate changes in biomass and ultimately radial size, while a customized Hybrid Coordinate Ocean Model (HYCOM) for the trajectory simulations was used. Their results showed that the trajectory of pelagic Sargassum was significantly influenced by inertia and the radial size of the rafts. In the inclusion of wind into studies has also been performed on an *ad-hoc* basis, where a fraction of wind velocity field is simply added to the ocean velocity field in terms of the velocity vectors (meridonal and zonal); for example 1%, 2%, 5% of the ocean velocity field has been used previously [@putman2018;@putman2020]. @putman2020 investigated the effect of including windage in kelp trajectory simulations and showed that including ad hoc windage factors greatly improved accuracy regarding entrainment or expulsion from eddies when compared to tracked Sargassum mats in the ocean. Langrangian trajectory modeling is clearly a useful tool; however, the various models available often assume the object to either be purely Lagrangian (no drag) or to be only influenced by windage (drag from the wind), which is obviously not the case for both biological and anthropogenic forms of marine debris in nature. The radial size in relation to windage included by @brooks2019 is an improvement but this approach does not consider hydrodynamic drag. Another example is work by @zhong2012 describing Sargassum distribution in relation to mesoscale and submesoscale eddies by simulating passive tracers. The authors used two model configurations, one with an idealised eddy-flow and another that simulates flow in the western Gulf of Mexico to advect passive tracers as a proxy for Sargassum. The authors showed that mesoscale and submesoscale features both play a role in accumulation zones for the virtual Sargassum and were able to explain some of the observed accumulation of Sargassum in the western Gulf of Mexico. Treating the virtual Sargassum particles as passive tracers may be the reason for the unexplained patterns during comparison of simulated and observed distributions. The authors do mention running simulations which included Stokes drag and inertia using a Maxey-Riley approach, and state that it made no significant difference to the overall results. However, the authors only used the density as a parameter for calculating the relevant forces by making the particles 10% lighter than the surrounding fluid. Additionally, the authors did not consider surface area or take hydrodynamic drag and windage into account, all of which are important influencers of trajectory.

Gyres, eddies, and converging surface currents have been identified as important in the accumulation and dispersal of plastics and other marine debris [@brach2018; @dasaro2018; @onink2019]. Investigations into the dispersal of marine debris have also included the possible effects of inertia in the accumulation of ocean drifter instruments in mesoscale features. For example, @beron2016 used a modelling approach to investigate the role of Ekman transport and inertia to explain the observed behavior of undrogued drifters in the ocean gyres. Undrogued drifter trajectory is affected by wind, unlike the drogue counterparts which makes the effects of wind negligible. The results showed that inertia, not Ekman transport, played a significant role in the observed drifter behaviour [@beron2016]. These authors also noted that the presence of other flotsam, along with the undrogued drifters in ocean gyres, suggests that aspects of inertia also play a role in accumulation of other marine debris [@beron2016]. The approach by @beron2016 has also been improved upon by @beron2019 but they also highlight the significant role of inertia in flotsam trajectory and accumulation.

This study aims to use a numerical approach to shed light on the role of hydrodynamic and direct wind drag on macroalgal drift trajectory using Lagrangian based trajectory simulations. The ‘behaviour’ of surface drifting kelp is investigated with specific focus on trajectories from a coastal environment to the open ocean. We achieved this by comparing the trajectory outputs of virtual passive particles (purely Lagrangian i.e no drag characteristics added, as a reference) with kelp particles (drag characteristics added) with surface area exposed to the various drag components.

# Methods

## Study site

The region of interest for this study was the Cape Peninsula in the Western Cape (Figure 1A). Kommetjie, located on the Atlantic side of the Cape Peninsula (Figure 1B), was chosen as a release site. The hydrodynamics in the study region are characterized by a complex interaction of wind, sea waves and ocean swell, and are also influenced by larger-scale regional oceanic processes. The larger processes in the region can be attributed to the Agulhas Current (AC) and the Benguela Current (BC) [@lutjeharms2000;@lutjeharms2007;@garzoli1996;@hutchings2009;@rubio2009]. The AC is a component of the South-west Indian Ocean sub-gyre, flowing predominantly south-westward following the continental shelf edge and eventually retroreflecting eastward back into the South Indian Ocean [@lutjeharms2000;@lutjeharms2007].

The region where the retroflection occurs is known as the Agulhas Retroflection and consists of high levels of mesoscale variability [@garzoli1996]. The variability results in shedding of rings and filaments known as the Agulhas Rings and have an average diameter of ~324km [@lutjeharms1988;@lutjeharms2006; @lutjeharms2007]. Agulhas Rings are the main contributors to the Agulhas leakage which is the transport of warm water from the Indian Ocean to the South Atlantic Ocean [@garzoli1996; @lutjeharms2006; @beal2015]. The Agulhas leakage is also a source of mesoscale variability within the Benguela region as anticyclonic rings are shed and drift northwestward along the southern Benguela slope region [@rubio2009]. The Benguela region ---known as the Benguela Upwelling System (BUS)--- is characterized by strong coastal upwelling which is primarily driven by equatorward winds [@shannon1996;@hardman2003;@veitch2018; @veitch2017]. The BUS starts at 27$^\circ$S and extends southward to 35$^\circ$S [@shannon1996]. The region is also characterized by large amounts of mesoscale variability in the form of eddys and filaments [@blanke2002; @blanke2005; @rubio2009]. Inshore, the BUS is dominated by a northwestward flow, known as the Benguela Jet, due to topographical steering, wind stress and interactions with passing Agulhas Rings and eddys [@stramma1989; @veitch2010; @veitch2017; @veitch2018; @hutchings2009]. The stability of the jet varies seasonally with more stable flows in winter; in January, the jet begins to split between Cape Point and Cape Columbine [@veitch2018; @blanke2005; @blanke2002] which results in a weak narrow inshore branch that flows along the shelf edge and a strong flowing north-westwards offshore branch. The area is also characterized by high upwelling activity driven by strong westerly winds [@hutchings2009]. These winds are most intense during Austral summer (December-March) which favours upwelling conditions in the region, and also act to reinforce the flow of the Benguela Jet which favours upwelling waters contributing to the across-shore density gradient [@ragoasha2019]. In Austral winter the wind direction is reversed which decreases upwelling favorable conditions, but the core of the Benguela Jet is maintained by the intrusion of warm Agulhas waters into the Benguela region over the shelf edge [@veitch2018; @hutchings2009; @ragoasha2019].

```{r map chp3, echo=FALSE, message=FALSE, warning=FALSE, strip.white=TRUE, fig.cap="Figure 1: Map of the study domain and the main features pertaining to transport in the region. In panel A, the map of the western portion of the South African coastline and the main currents is presented. The resulting eddies and filaments forming from the Agulhas retroflection are indicated, and the position of the study area is drawn in as a black rectangle. Details of the study area are given in panel B. Here the splitting of the Benguela Jet, eddy features, and the general flow within the study domain are presented. The release site of simulated kelp ‘particles’ is represented by a black triangle."}

@@

## Numerical models

Virtual particles were released with no drag effects from the study site to act as a reference for comparison to simulations with varying combinations of hydrodynamic and wind drag and are referred to hereafter as ‘passive particles.’ All particles that have any form of drag exposure are hereafter referred to as ‘kelp particles.’ All simulations have the same release location. The release site, Kommetjie, was in an area of a known kelp population, which a previous study by @coppin2020 collected morphological data from. Due to the resolution of the underlying ocean model (see Ocean and wind model section) nearshore processes are not resolved well. Therefore, the release site was placed in the adjacent grid cell in order to better represent offshore transport and to avoid model limitations with regards to nearshore transport. Since the aim of this investigation was to investigate the relative surface behaviour of virtual kelp particles, only the surface layer of the underpinning hydrodynamic models were used. It should also be noted that the Copernicus model (see Ocean and wind model section) used velocity vectors over some of the land covering the coastal domain.

The time period used in all simulations were Austral summer, autumn, winter, and spring of 2018. These time periods were chosen based on variable mesoscale activities and stability of the Benguela Jet that occurs throughout the year. For example, splitting of the Benguela Jet occurs in January, which also coincides with a period of higher wind intensity, upwelling along the coast and mesoscale activity while in winter the Benguela Jet begins to stabilize until summer [@rubio2009; @Veitch2019; @ragoasha2019]. Simulations were executed for the duration of each season, with 3-hourly timesteps and 3-hourly outputs using a fourth-order Runge-Kutta advection scheme for both passive and kelp particles. Drag components for kelp particles were incorporated by the modifiable component of the OceanParcels advection scheme [@North2009; @vanSebille2018].

To test the different magnitudes that hydrodynamic and wind drag might play in the overall trajectory, simulations were run with degrees of drag exposure scenarios based on personal observations of drifting kelp. This acts as a proxy for buoyancy, which determines the surface area exposed to hydrodynamic and wind drag. These estimates were expressed as percentages which were used to calculate hydrodynamic and wind drag for the applicable simulation. The scenarios used in this study were 100%  and 85% for hydrodynamic and Stokes drag and 0% and 15% for wind.  Simulations were separated into fully submerged (100% hydrodynamic drag and 0% wind drag, see Figure 2A) and partially submerged (85% hydrodynamic drag, and 15% wind drag, See Figure 2C-D) categories. For each of the categories, particles were advected with different combinations of flow-fields (currents, currents-Stokes drift, currents-wind, currents-wind-Stokes drift).  A similar approach has been used in past research investigating the role of windage in distribution of pelagic Sargassum rafts by @putman2018 and @putman2020 in the form of ad hoc windage estimates; although the authors did not include hydrodynamic drag unlike the current study.

To determine if vortices and eddies play a role in kelp accumulation on the surface, a separate set of simulations were performed. Unlike the previous simulations, these simulations released 10 particles over the latitudinal extent of the study domain from the study site, equidistant from each other. A summary of the simulations executed are provided in table 2.

@@

The model is based on parameterizations that have been used for previous studies investigating the dispersal patterns of various flotsam by implementing custom parameterisations of Lagrangian trajectory models [@furnans2008; @beron2016; @hart2018; @brooks2019; @putman2020; @jutzeler2020]. Once dislodged, kelp can accumulate on the surface and become entangled to form a dense mat known as a kelp-raft [@hobday2000a; @smith2002]. Kelp rafts are not a common feature around the coast of the Cape Peninsula; rather, solitary kelp is (pers. obs). Therefore, the trajectory of only solitary *E. maxima* individuals will be considered. All simulations in this study used the OceanParcels software package (version 2.0), which is an open source, customizable Lagrangian Trajectory tool [@delandmeter2019]. OceanParcels calculates the advection of particle trajectory with a Lagrangian method [@delandmeter2019], which can be described by the equation:

$$X(t+\Delta t) = X(t) + \int_{t}^{t+\Delta t} v(x, \tau)d \tau + \Delta X_b(t)$$
